---
title: "baseR_part2"
author: "Pageneck Chikondowa"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Reading the object
```{r}
my_matrix <- readRDS("my_day2matrix.rds")
print(my_matrix)
```
Adding row sums and means to this dataset
```{r}
data_matrix <- cbind(my_matrix, rowSums=rowSums(my_matrix), rowMeans=rowMeans(my_matrix))
data_matrix
```


```{r}
data_matrix2 <- rbind(data_matrix, c(colSums(my_matrix), NA, NA), c(colMeans(my_matrix),NA,NA))
data_matrix2
```

Adding the row names to the matrix

```{r}
rownames(data_matrix2) <- (c("sample1","sample2","sample3","sample4","sample5","sample6","sample7","sample8","sample9","sample10","sample11","sample12","sample13","sample14","sample15","sample16","sample17","sample18","sample19","sample20","Sum","Mean"))
data_matrix2

```

Loading the ToothGrowth dataset

```{r}
data(ToothGrowth)
head(ToothGrowth)
```

```{r}
# Descriptive summary statistics for the Toothbrush dataset: 
summary(ToothGrowth)
```

```{r}
#The average tooth length is: 
mean(ToothGrowth$len)
```

```{r}
print(paste("minimum teeth length: ", min(ToothGrowth$len)))
print(paste("maximum teeth length: ", max(ToothGrowth$len)))

```

```{r}
hypotenuse  <- function(side1,side2){
  hypo <- sqrt(side1**2 + side2**2)
  return(hypo)
}
```


### functions
```{r}
hypotenuse(3,4)
```

### lapply() and sapply()

Exercise 7 - apply

part 1
```{r}
myvector <- c(1:10)
```

A function that squares numbers
```{r}
sqaure <- function(x){
  sq <- x^2
  return(sq)
}

```

using lappy
```{r}
lapply(myvector, sqaure)

```

using sapply
```{r}
sapply(myvector, sqaure)
```

Part 2

creating a list with 4 items
```{r}
my_list <- list(myvector, third,a, c(T,F,F,T))
```


using the lappy function
```{r}
lapply(my_list,sum)
```
using the sapply function

```{r}
sapply(my_list, sum)
```   

using the sapply

definign the repeating function

```{r}
sapply(my_list,rep,each = 3)

```

### Loops 





